MISUSES AND INACCURACIES OF PAGEL'S LAMBDA ML ESTIMATION METHODS

### 2020-04-27 Manuscript/Manuscript-3.Rmd
vsn 3

Evolutionary biologists commonly seek to describe the relative strength of phylogenetic signal in phenotypic traits, to determine the extent to which shared evolutionary history has influenced trait covariation among taxa. This is often accomplished by interpreting empirical estimates of $\lambda$; with smaller values signifying 'weak' phylogenetic signal, while larger values are interpreted as 'strong' phylogenetic signal [e.g., @DeMeester2019; @Pintanel2019;@Su2019]. Other approaches for interpreting $\lambda$ are more statistical, through the use of confidence intervals [@Vandelook2019] or likelihood ratio tests that compare the observed model fit to that obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019]. Likewise, qualitative comparisons of $\lambda$ across multiple phenotypic traits have also been used to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019]. Indeed, it seems intuitive to interpret the strength of phylogenetic signal in this manner, as $\lambda$ is a parameter on a bounded scale ($0\to1$) for which interpretation of its extremal points are understood ($\lambda=0$ represents no phylogenetic signal, while $\lambda=1$ is phylogenetic signal as expected under Brownian motion). However, equating values of $\lambda$ directly to the strength of phylogenetic signal presumes two important statistical properties that have not been fully explored. \hfill\break

First, it presumes that values of $\lambda$ can be precisely estimated, as biological inferences regarding the strength of phylogenetic signal depend on high accuracy in its estimation. Therefore, understanding the precision in estimating $\lambda$ is paramount. One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation (i.e., greater precision) when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$). From this observation it was concluded that insufficient data (i.e., the number of species) was the underlying cause of the increased variation across parameter estimates [@Boettiger_et_al2012]. Indeed, such a pattern is common with statistical estimators, as summary statistics and parameters are often more precise at greater sample sizes [@Cohen1988]. However, this conclusion also assumes that the precision of $\lambda$ remains constant across its range ($\lambda = 0 \to 1$); an assumption that to date, has not been verified. Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we still lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets. \hfill\break

@@

In this study, we evaluate the precision of Pagel's $\lambda$ in estimating known levels of phylogenetic signal in phenotypic data. We use computer simulations with differing numbers of species, differently shaped phylogenies, and differing input levels of phylogenetic signal, to explore the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal, and under what circumstances. We find that while PGLS parameters (e.g., $\beta$) are accurately estimated with the inclusion of phylogenetic signal, estimates of $\lambda$ are not. We also find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that the precision in estimating $\lambda$ is not constant across the range of input signal, with decreased precision when phylogenetic signal is of intermediate strength. Additionally, the same $\lambda_{est}$ may be obtained from datasets containing vastly different input levels of phylogenetic signal. Thus, $\lambda$ is not a reliable estimate of the strength of phylogenetic signal in phenotypic data. We subsequently derive a standardized effect size for measuring the strength of phylogenetic signal in phenotypic datasets, and apply the concept to two common measures of phylogenetic signal: $\lambda$ and $Kappa$. Through simulations across a wide range of conditions, we find that the precision of effect sizes based on $\lambda$ ($Z_{\lambda}$) are less reliable than that those based on $Kappa$ ($Z_K$), implying that $Z_K$ is a more robust effect size measure. Additionally, we propose a two-sample test statistic that may be used to compare the strength of phylogenetic signal among datasets, and provide an empirical example to demonstrate its use. We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often inaccurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure is compromised. By contrast, effect sizes obtained from $Kappa$ hold promise for characterizing phylogenetic signal, and for comparing the strength of phylogenetic signal across datasets.

@@

*Results.* We found that the precision of $\lambda_{est}$ varied widely across simulation conditions. Predictably, precision improved as the number of species increased (Figure 1). This confirmed earlier findings of Boettiger et al. [-@Boettiger_et_al2012], and adhered to parametric statistical theory. However, in many cases the set of $\lambda_{est}$ spanned nearly the entire range of possible values (e.g., $n=32$; $\lambda_{in}=0.5$: $\lambda_{est}= 0.0\to 0.985$), revealing that estimates of $\lambda$ were not a reliable indicator of input phylogenetic signal. Importantly, the precision of $\lambda_{est}$ was not uniform across all levels of phylogenetic signal, with the worst precision at intermediate levels of signal ($\lambda_{in}\approx0.5$), and improved precision as input levels approached the extremes of its range (i.e., $\lambda_{in}\rightarrow0$ & $\lambda_{in}\rightarrow1$). Thus, estimates of $\lambda$ were least reflective of the true input signal at intermediate values. Additionally, even at large levels of species richness, we found that the range of $\lambda_{est}$ still encompassed a substantial portion of possible values (e.g., $n=512$; $\lambda_{in}=0.5$: $\lambda_{est} = 0.32\to 0.68$). Likewise, the same $\lambda_{est}$ could be obtained from datasets containing vastly different input levels of phylogenetic signal (e.g., $n=512$; $\lambda_{est} = 0.5$; $\lambda_{in} = 0.25 \to 0.65$). Results were similar when $\lambda$ was co-estimated with regression parameters in PGLS regression (Figure 2). Here, regression parameters ($\beta$) were accurately estimated, confirming earlier findings of Revell 2010 [-@Revell2010] (see Supporting Information). However, estimates of phylogenetic signal were not, and the spread of $\lambda_{est}$ was even broader than that observed when $\lambda$ was estimated for only the dependent variable. Taken together, these findings reveal that $\lambda_{est}$ does not precisely characterize observed levels of phylogenetic signal in phenotypic datasets, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ may be highly inaccurate. \hfill\break

@@

@@

Adams and Collyer [-@AdamsCollyer2019] recently suggested that the strength of phylogenetic signal could be represented an effect size, obtained from $Kappa$ and its empirical sampling distribution from permutation. Here we formalize that suggestion, and find an effect size as:

@@

Similarly, an effect size based on $\lambda$ could be envisioned as:

\begin{align}
    Z_{\lambda}=\frac{\lambda_{obs}-0}{\hat\sigma_\lambda}
\end{align}

Note that under the null hypothesis, $E(\lambda)=0$, a no phylogenetic signal is expected under this condition [@Freckleton_et_al2002]. \hfill\break

To evaluate the utility of $Z_K$ and $Z_\lambda$ we calculated both effect sizes for the simulated datasets generated above, and summarized the precision of each using its variance ($\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$). Results are found in Figure 3. Here two things are evident. First, estimates of $Z_K$ track the input phylogenetic signal in a more linear fashion than to estimates of $Z_\lambda$. Second, the precision of $Z_K$ is considerably more stable, as coefficients of variation for the set of $\sigma^2_{Z_K}$ across input levels of phylogenetic signal for the same sample size was an order of magnitude smaller than that observed for $\sigma^2_{Z_\lambda}$ (Figure 3). This implies that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$ as compared with $Z_\lambda$. \hfill\break

[insert Figure 3 here]

@@

Once the magnitude of phylogenetic signal is characterized using $Z_K$, it may be of interest to compare such measures across datasets. This is useful, for instance, to determine whether the strength of phylogenetic signal is greater in one phenotypic trait as compared with another. As with other effect sizes derived from permutation distributions [e.g., @AdamsCollyer2016; @AdamsCollyer2019b], a two-sample test statistic may be found as:

@@

where $K_1$, $K_2$, $\hat\mu_{K_1}$, $\hat\mu_{K_2}$, $\hat\sigma_{K_1}$, and $\hat\sigma_{K_2}$ are as defined above for equation 2. Estimates of significance of $\hat{Z}_{12}$ may be obtained from a standard normal distribution. As with other two-sample tests, $\hat{Z}_{12}$ is typically considered a two-tailed test, however directional (one-tailed) tests may be specified should the empirical situation require it [see @AdamsCollyer2016; @AdamsCollyer2019b].

@@

## *Literature Survey... WHERE DOES THIS GO???*

To determine how Pagel's $\lambda$ is commonly utilized in empirical studies, we conducted a literature survey. From Google.scholar we obtained a list of all papers published in 2019 that used $\lambda$; resulting in 341 studies. For each study, we extracted all $\lambda_{est}$, the size of the phylogeny ($n$), and noted whether authors reported confidence intervals or performed significance tests assessing difference of $\lambda_{est}$ from either zero or one. We also noted whether biological interpretations based on $\lambda_{est}$ were made, and for studies that reported more than one $\lambda_{est}$, we also noted whether these were compared in some manner, and whether such comparisons were accompanied with statistical tests between $\lambda_{est}$.

*Results*. We found 182 manuscripts from 2019 that estimated and reported Pagel's lambda values using PGLS methods. These papers averaged 8.527 lambda values, ranging from a single lambda estimate up to 71 estimated lambdas. Almost exactly half of the published lambda estimates were either below 0.05 (25.32%) or above 0.9 (24.74%; Figure 3). 73.32% of the published lambdas were estimated using phylogenies with fewer than 200 tips, and 348 lambda estimates (8.57% of all published estimates) came from phylogenies with fewer than 30 tips.  \hfill\break

[insert Figure 4 here] \hfill\break

Many of the reviewed manuscripts liberally interpreted the magnitude of the estimate lambda, using phrases such as "strong" or "weak" phylogentic signal when statistically, all that was clear was a difference between the estimated lambda and 0 or 1 respectively. We estimated that about 20.49% of the manuscripts revealed some sort of biological interpretation of the magnitude of estimated phylogenetic signal that overreached the statistical findings. We also identified seven manuscripts as having inappropriately interpreted differences in lambda values, indicating that some traits had stronger or weaker signal than other traits without the appropriate statistical tests. \hfill\break

As is evidenced by macroevolutionary papers published in 2019 papers, Pagel's lambda estimation methods are often misused and over-interpretted. Despite the urging of Boettiger and colleagues to publish confidence intervals with all lambda parameter estimates, only 18% of papers published in 2019 do so.

# Discussion

@@

\textbf{Figure 1}. Precision of Pagel's $\lambda$ across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes. As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal. \hfill\break

\textbf{Figure 2}. Precision of Pagel's $\lambda$ when incorporated in phylogenetic regression ($Y\sim X$), across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes. As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal. \hfill\break

\textbf{Figure 3}.  Variation in estimates of phylogenetic signal across input levels of phylogenetic signal. (A) Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), (B) Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), (C) Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species. (D) Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.    \hfill\break

\textbf{Figure 4}. Frequency of estimated lambda values published in manuscripts in 2019. The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa.

@@

\singlespacing \textbf{Figure 1}. Precision of Pagel's $\lambda$ across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes. As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal.

\newpage
```{r, echo = FALSE, out.width="95%"}
include_graphics("Fig2.png")

@@

\singlespacing \textbf{Figure 2}. Precision of Pagel's $\lambda$ when incorporated in phylogenetic regression ($Y\sim X$), across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes. As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal.

@@

\singlespacing \textbf{Figure 3}.  Variation in estimates of phylogenetic signal across input levels of phylogenetic signal. (A) Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), (B) Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), (C) Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species. (D) Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.

@@

\singlespacing \textbf{Figure 4}.Frequency of estimated lambda values published in manuscripts in 2019. The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa.

>>>>

Evolutionary biologists commonly seek to describe the relative strength of phylogenetic signal in phenotypic traits, to determine the extent to which shared evolutionary history has influenced trait covariation among taxa. This is often accomplished by interpreting empirical estimates of $\lambda$; with smaller values signifying 'weak' phylogenetic signal, while larger values are interpreted as 'strong' phylogenetic signal [e.g., @DeMeester2019; @Pintanel2019;@Su2019]. Other approaches for interpreting $\lambda$ are more statistical. For instance, some have evaluated whether the observed $\lambda$ differs from some expected value through the use of confidence intervals [@Vandelook2019] or by performing likelihood ratio tests that compare the observed model fit to that obtained when $\lambda=0$ or $\lambda=1$ [@Freckleton_et_al2002; @Cooper2010; @Bose2019]. Additionally, qualitative comparisons of $\lambda$ estimates obtained from multiple phenotypic traits have been used to infer whether the strength of phylogenetic signal is greater in one trait as compared to another [e.g., @Liu2019; @Bai2019]. Indeed, statements regarding the strength of phylogenetic signal based on $\lambda$ are rather common in the evolutionary literature. We conducted a literature survey in Google.scholar and found that of the 204 papers published in 2019 that estimated and reported Pagel's $\lambda$, 40% interpreted the strength of phylogenetic signal for at least one phenotypic trait. Additionally, because nearly half of the 1,572 $\lambda$ values reported were near the limits of the parameter (Figure 1), this percentage is even higher, as biological interpretation of phylogenetic signal at the limits of $\lambda$ are known. \hfill\break

It seems intuitive to interpret the strength of phylogenetic signal based on the value of $\lambda$, as $\lambda$ is a parameter on a bounded scale ($0\to1$) for which interpretation of its extremal points are understood. Specifically, $\lambda=0$ represents no phylogenetic signal, while $\lambda=1$ is phylogenetic signal as expected under Brownian motion. However, equating values of $\lambda$ directly to the strength of phylogenetic signal presumes two important statistical properties that have not been fully explored. First, it presumes that values of $\lambda$ can be precisely estimated, as biological inferences regarding the strength of phylogenetic signal depend on high accuracy in its estimation. Therefore, understanding the precision in estimating $\lambda$ is paramount. One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation (i.e., greater precision) when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$). From this observation it was concluded that insufficient data (i.e., the number of species) was the underlying cause of the increased variation across parameter estimates [@Boettiger_et_al2012]. Indeed, such a pattern is common with statistical estimators, as summary statistics and parameters are often more precise at greater sample sizes [@Cohen1988]. However, this conclusion also assumes that the precision of $\lambda$ remains constant across its range ($\lambda = 0 \to 1$); an assumption that to date, has not been verified. Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we still lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets. \hfill\break

@@

In this study, we evaluate the precision of Pagel's $\lambda$ in estimating known levels of phylogenetic signal in phenotypic data. We use computer simulations with differing numbers of species, differently shaped phylogenies, and differing input levels of phylogenetic signal, to explore the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal, and under what circumstances. We find that while PGLS parameters (e.g., $\beta$) are accurately estimated with the inclusion of phylogenetic signal, estimates of $\lambda$ are not. We also find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that the precision in estimating $\lambda$ is not constant across the range of input signal, with decreased precision when phylogenetic signal is of intermediate strength. Additionally, the same estimated values of $\lambda$ may be obtained from datasets containing vastly different input levels of phylogenetic signal. Thus, $\lambda$ is not a reliable estimate of the strength of phylogenetic signal in phenotypic data. We subsequently derive a standardized effect size for measuring the strength of phylogenetic signal in phenotypic datasets, and apply the concept to two common measures of phylogenetic signal: $\lambda$ and $Kappa$. Through simulations across a wide range of conditions, we find that the precision of effect sizes based on $\lambda$ ($Z_{\lambda}$) are less reliable than that those based on $Kappa$ ($Z_K$), implying that $Z_K$ is a more robust effect size measure. Additionally, we propose a two-sample test statistic that may be used to compare the strength of phylogenetic signal among datasets, and provide an empirical example to demonstrate its use. We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often inaccurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure is compromised. By contrast, effect sizes obtained from $Kappa$ hold promise for characterizing phylogenetic signal, and for comparing the strength of phylogenetic signal across datasets.

@@

*Results.* We found that the precision of $\lambda_{est}$ varied widely across simulation conditions. Predictably, precision improved as the number of species increased (Figure 2). This confirmed earlier findings of Boettiger et al. [-@Boettiger_et_al2012], and adhered to parametric statistical theory. However, in many cases the set of $\lambda_{est}$ spanned nearly the entire range of possible values (e.g., $n=32$; $\lambda_{in}=0.5$: $\lambda_{est}= 0.0\to 0.985$), revealing that estimates of $\lambda$ were not a reliable indicator of input phylogenetic signal. Importantly, the precision of $\lambda_{est}$ was not uniform across all levels of phylogenetic signal, with the worst precision at intermediate levels of signal ($\lambda_{in}\approx0.5$), and improved precision as input levels approached the extremes of its range (i.e., $\lambda_{in}\rightarrow0$ & $\lambda_{in}\rightarrow1$). Thus, estimates of $\lambda$ were least reflective of the true input signal at intermediate values. Additionally, even at large levels of species richness, we found that the range of $\lambda_{est}$ still encompassed a substantial portion of possible values (e.g., $n=512$; $\lambda_{in}=0.5$: $\lambda_{est} = 0.32\to 0.68$). Likewise, the same $\lambda_{est}$ could be obtained from datasets containing vastly different input levels of phylogenetic signal (e.g., $n=512$; $\lambda_{est} = 0.5$; $\lambda_{in} = 0.25 \to 0.65$). Note that these findings were particularly unsettling when considered in light of our literature survey. Over one quarter of the $\lambda$ estimates obtained in empirical studies (421 of 1,572) were between $\lambda=0.25$ and $\lambda=0.75$ (Figure 1). Of those, 30% were interpreted directly in terms of the strength of phylogenetic signal that they represented (e.g., small, intermediate, strong signal). \hfill\break

Finally, when $\lambda$ was co-estimated with regression parameters in PGLS regression, the results of our simulations were quite similar. Here, regression parameters ($\beta$) were accurately estimated, confirming earlier findings of Revell [-@Revell2010] (see Supporting Information). Estimates of phylogenetic signal were less precise, and the spread of $\lambda_{est}$ was similar to that observed when $\lambda$ was estimated for only the dependent variable (Figure 3). Taken together, these findings reveal that $\lambda_{est}$ does not precisely characterize observed levels of phylogenetic signal in phenotypic datasets, and that biological interpretations of the strength of phylogenetic signal based on $\lambda$ may be highly inaccurate. \hfill\break

@@

[insert Figure 3 here] \hfill\break

@@

Adams and Collyer [-@AdamsCollyer2019] recently suggested that the strength of phylogenetic signal could be represented as an effect size, based on the $Kappa$ statistic and its empirical sampling distribution from permutation. Here we formalize that suggestion, resulting in an effect size of:

@@

where $K_{obs}$ is the observed phylogenetic signal, and $\hat\mu_K$ and $\hat\sigma_K$ are the mean and standard deviation of the empirical sampling distribution of $Kappa$ obtained via permutation. Similarly, an effect size based on $\lambda$ could be envisioned as:

\begin{align}
    Z_{\lambda}=\frac{\lambda_{obs}-0}{\hat\sigma_\lambda}.
\end{align}

In this case, $\lambda_{obs}$ and $\hat\sigma_\lambda$ are empirically derived using maximum likelihood. Note that under the null hypothesis no phylogenetic signal is expected [@Freckleton_et_al2002], and thus $E(\lambda)=0$ under this condition. \hfill\break

To evaluate the utility of $Z_K$ and $Z_\lambda$ we calculated both effect sizes for the simulated datasets generated above, and summarized the precision of each using its variance ($\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$). Results are found in Figure 4. Here two things are evident. First, estimates of $Z_K$ track the input phylogenetic signal in a more linear fashion than do estimates of $Z_\lambda$ (Figure 4A,B). Thus, actual changes in the strength of phylogenetic signal are reflected more evenly in the corresponding values of the effect size $Z_K$. Second, the precision of $Z_K$ is considerably more stable as compared with $Z_\lambda$. This may be seen by calculating the coefficients of variation for the set of precision estimates (i.e., $\sigma^2_{Z_K}$ and $\sigma^2_{Z_\lambda}$) across input levels of phylogenetic signal.  Here coefficients of variation in the precision of $Z_K$ were an order of magnitude smaller for than for $Z_\lambda$ (Figure 4C,D). This implied that estimates of the strength of phylogenetic signal were more reliable and robust when using $Z_K$ as compared with $Z_\lambda$. \hfill\break

[insert Figure 4 here]

@@

Once the magnitude of phylogenetic signal is characterized using $Z_K$, one may wish to compare such measures across datasets, to determine whether the strength of phylogenetic signal is significantly greater in one phenotypic trait as compared with another. As with other effect sizes derived from permutation distributions [e.g., @AdamsCollyer2016; @AdamsCollyer2019b], a two-sample test statistic may be found as:

@@

where $K_1$, $K_2$, $\hat\mu_{K_1}$, $\hat\mu_{K_2}$, $\hat\sigma_{K_1}$, and $\hat\sigma_{K_2}$ are as defined above for equation 2. Estimates of significance of $\hat{Z}_{12}$ may be obtained from a standard normal distribution. Typically, $\hat{Z}_{12}$ is considered a two-tailed test, however directional (one-tailed) tests may be specified should the empirical situation require it [see @AdamsCollyer2016; @AdamsCollyer2019b]. \hfill\break

One important advantage of the approach advocated here is that the resulting effect sizes ($Z_K$) are dimensionless, as the units of measurement cancel out during their calculation [**S&R??**]. As a consequence, the effect sizes represent the strength of phylogenetic signal on a common and comparable scale, measured in standard deviation units, regardless of the initial units and scale of the original phenotypic variables under investigation.

###
Conceptually, this is akin to the conversion of data to their standard normal deviates prior to their use in multivariate analysis, when the original variables are expressed in different units or scales [see ]. 
###

####

mathematical transformation or standardization of the trait dimensions, so
the variables are expressed in similar units
n the variables in a multivariate data set are on different scales,

Everitt, Brian; Hothorn, Torsten J (2011), An Introduction to Applied Multivariate Analysis with R, Springer,

Also, Legendre... Adams and Collyer. 
####

Because of this standardization, the strength of phylogenetic signal across datasets has meaning

something about comparisons having meaning...

As such, characterizing the strength of phylogenetic signal with $Z_K$ facilitates a formal statistical comparison of values across continuous traits of different units, scale, or even dimensionality (e.g., univariate traits using $Kappa$, and multivariate traits using $K_{mult}$).

@@

@@

\textbf{Figure 1}. Frequency distribution of $\lambda$ estimates published in 2019. The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa. \hfill\break

\textbf{Figure 2}. Precision of Pagel's $\lambda$ across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes. As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal. \hfill\break

\textbf{Figure 3}. Precision of Pagel's $\lambda$ when incorporated in phylogenetic regression ($Y\sim X$), across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes. As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal. \hfill\break

\textbf{Figure 4}.  Variation in estimates of phylogenetic signal across input levels of phylogenetic signal. (A) Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), (B) Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), (C) Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species. (D) Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.

@@

\singlespacing \textbf{Figure 1}.  Frequency distribution of $\lambda$ estimates published in 2019. The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa.

```{r, echo = FALSE, out.width="95%"}
include_graphics("Fig2.png")

@@

\singlespacing \textbf{Figure 2}. Precision of Pagel's $\lambda$ across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes. As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal.

@@

\singlespacing \textbf{Figure 3}. Precision of Pagel's $\lambda$ when incorporated in phylogenetic regression ($Y\sim X$), across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes. As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal.

@@

\singlespacing \textbf{Figure 4}.  Variation in estimates of phylogenetic signal across input levels of phylogenetic signal. (A) Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), (B) Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), (C) Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species. (D) Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.
